---
title: "Medical expenses prediction using regression analysis in R"
#author: "Melvin Galera"
#date: "2024-03-06"
output: 
  github_document:
    toc: false
---

```{r setup, include=FALSE}
# Global code chunk options; adjust individual codes as required
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 12)

# Load data
insurance_df <- read.csv("./data/insurance.csv")
insurance_df_og <- insurance_df

# Load libraries
library(tidyverse)
library(cowplot)
library(kableExtra)
library(scales)
library(ggcorrplot)
library(gridExtra)
library(ggpubr)
```

<br>

### I. Project objective

Health insurance is important because it provides protection from disruptive medical expenses which might result from an illness, disability or accident. Insurance companies, based on the health insurance coverage, can reimburse the insured for medical expenses incurred or can pay the care provider directly. It is important for insurers to be able to forecast medical expenses so they could accurately set yearly premiums that their beneficiaries have to pay to avail health insurance.  

Hence, this project was aimed to develop a predictive model for estimating medical expenses based on given predictor variables using linear regression method. In this case, predictor variables were based on patient characteristics data.

<br>

### II. Exploratory Data Analysis

**A. Data Collection**

For this analysis, a simulated dataset containing medical expenses for patients in the United States obtained from Kaggle was used. This dataset was created for the book "Machine Learning with R" (by Brett Lantz, 2013) using demographic statistics from the US Census Bureau. The sourced dataset has 1338 records and 7 variables indicating patient characteristics including the total medical expenses (referred as `charges`). The variables are:

  * `age`       : the person's age
  * `sex`       : the person's gender, either male or female
  * `bmi`       : the person's body mass index which indicates how over or under-weight the person is relative to their height
  * `children`  : the number of children (or dependents) in the person's family
  * `smoker`    : indicates if the person is a regular smoker or not ('yes' or 'no')
  * `region`    : the US geographic region where the person resides ('southwest', 'southeast', 'northwest', 'northeast')
  * `charges`   : the amount of medical expenses (in US dollars) charged to the insurance plan for the calendar year

<br>

**B. Data Exploration**

An initial look at the structure of the loaded dataset (`insurance_df`)shows:

```{r, echo= FALSE}
str(insurance_df)
```
The first 10 records of the `insurance_df`:

```{r}
insurance_df %>% head(10)
```

<br>

To perform EDA on `insurance_df`, univariate distribution of variables, as well as bivariate and multivariate relationships among the variables were investigated.

<br>

**1. Univariate Analysis**

The figures below shows the distribution plots for medical expenses and patients BMI. The summary statistics are also described.



```{r, fig.align='center', fig.width=15, fig.height=6}
# create plot for medical expenses
p1_charges <- ggplot(data = insurance_df, aes(x = charges)) +
  geom_histogram(binwidth= 5000, color = "dodgerblue1", fill = "dodgerblue3") +
  scale_x_continuous(labels = dollar_format())+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16)) +
  labs(x = "Charges",
       y = "Count",
       title = "Charges distribution")

# create plot for bmi
p2_bmi <- ggplot(data = insurance_df, aes(x = bmi)) +
  geom_histogram(binwidth = 3, color = "goldenrod1", fill = "goldenrod3") +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16)) +
  labs(x = "BMI",
       y = "Count",
       title = "BMI distribution")

# combined plots
ggarrange(p1_charges, p2_bmi, ncol=2) #, labels = c ("A", "B", "C"))
```
<br>

*Summary statistics for medical `charges`:*

```{r}
# get summary statistics for medical expenses
summary(insurance_df$charges)
```

*Summary statistics for patients `bmi`:*

```{r}
# get summary statistics for BMI
summary(insurance_df$bmi)
```

<br>

The figures below shows the distribution bar plots for the patients' age group, number of children, region, sex, and smoker status. The summary statistics are also described.

```{r}
# Create age group
age_breaks <- c(1, 20, 30, 40, 50, 60, 70)

age_group_tags <- c("<20", "20-29", "30-39", "40-49", "50-59", "60+")

insurance_df$age_group <- cut(insurance_df$age, breaks = age_breaks,
                              include.lowest = TRUE,
                              right = FALSE,
                              labels = age_group_tags)

```


```{r, include = FALSE}
# Sex
is.factor(insurance_df$sex)
# Output is False

#Convert to a factor
insurance_df$sex <- factor(insurance_df$sex)

#check for missing values
sum(is.na(insurance_df$sex))
#output is zero (0)
```

```{r, include = FALSE}
# Smoker
is.factor(insurance_df$smoker)
# Output is False

#Convert to a factor
insurance_df$smoker <- factor(insurance_df$smoker)

#check for missing values
sum(is.na(insurance_df$smoker))
#output is zero (0)
```

```{r, include = FALSE}
# Region
is.factor(insurance_df$region)
# Output is False

#Convert to a factor
insurance_df$region <- factor(insurance_df$region)

#check for missing values
sum(is.na(insurance_df$region))
#output is zero (0)
```

```{r, include = FALSE}
# Number of children
is.factor(insurance_df$children)
# Output is False 

#Convert to a factor
insurance_df$children_num <- factor(insurance_df$children)

#check for missing values
sum(is.na(insurance_df$children_num))
#output is zero (0)
```

<br> 

```{r, fig.align='center', fig.width=15, fig.height=6}
# create a bar plot for age group
p3_age_group <- ggplot(data = insurance_df, aes(x = age_group, fill= age_group)) +
  geom_bar(width = 0.7, color = "dodgerblue1", fill = "dodgerblue3") +
  geom_text(data = . %>% 
              group_by(age_group) %>% 
              tally() %>% 
              mutate(prop = round(n/sum(n),3)) %>% 
              ungroup(),
            aes(y = n, label = scales::percent(prop)),
            position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16)) +
  labs(x = "Age Group",
       y = "Count",
       title = "Age distribution")


# create a bar plot for number of children (children_num)
p4_children_num <- ggplot(data = insurance_df, aes(x= children_num, fill = children_num)) +
  geom_bar(width = 0.7, color = "goldenrod1", fill = "goldenrod3")+
  geom_text(data = . %>% 
              group_by(children_num) %>% 
              tally() %>% 
              mutate(prop = round(n/sum(n),3)) %>% 
              ungroup(),
            aes(y = n, label = scales::percent(prop)),
            position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  labs(x= "Number of children",
       y = "Count",
       title = "Number of children distribution")+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16)) 
 

# combined plots
ggarrange(p3_age_group, p4_children_num, ncol=2) 
```

<br>

```{r, fig.align='center', fig.width=14, fig.height=6}
# create a bar plot for sex
p5_sex <- ggplot(data = insurance_df, aes(x= sex, fill = sex, color= sex)) +
  geom_bar(width = 0.7)+
  geom_text(data = . %>% 
              group_by(sex) %>% 
              tally() %>% 
              mutate(prop = round(n/sum(n),3)) %>% 
              ungroup(),
            aes(y = n, label = scales::percent(prop)),
            position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  scale_fill_manual(values = c("dodgerblue3", "goldenrod3"))+
  scale_color_manual(values = c("dodgerblue1", "goldenrod1"))+
  labs(x= "Sex",
       y = "Count",
       title = "Sex distribution")+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "none")

# create a bar plot for smoker
p6_smoker <- ggplot(data = insurance_df, aes(x= smoker, fill = smoker, color= smoker)) +
  geom_bar(width = 0.7)+
  geom_text(data = . %>% 
              group_by(smoker) %>% 
              tally() %>% 
              mutate(prop = round(n/sum(n),3)) %>% 
              ungroup(),
            aes(y = n, label = scales::percent(prop)),
            position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  scale_fill_manual(values = c("dodgerblue3", "goldenrod3"))+
  scale_color_manual(values = c("dodgerblue1", "goldenrod1"))+
  labs(x= "Smoker status",
       y = "Count",
       title = "Smoker status distribution")+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16)) +
  theme(legend.position = "none")


# create a bar plot for region
p7_region <- ggplot(data = insurance_df, aes(x= region, fill = region)) +
  geom_bar(width = 0.7, color = "dodgerblue1", fill = "dodgerblue3")+
  geom_text(data = . %>% 
              group_by(region) %>% 
              tally() %>% 
              mutate(prop = round(n/sum(n),3)) %>% 
              ungroup(),
            aes(y = n, label = scales::percent(prop)),
            position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  labs(x= "Region",
       y = "Count",
       title = "Region distribution")+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))


# combined plots
ggarrange( p7_region, p5_sex, p6_smoker, ncol=3, widths = c(1.3, 0.85, 0.85)) 
```
<br>

*Summary statistics for patients' `age`:*

```{r, fig.align='center', fig.width=8, fig.height=6}
# get summary statistics for age
summary(insurance_df$age)
```

*Distribution of number of `children` based on count:*
```{r}
table(insurance_df$children_num)
```


*Distribution of `region` residence based on count:*
```{r}
table(insurance_df$region)
```

*Distribution of `sex` based on count:*
```{r}
table(insurance_df$sex)
```

*Distribution of `smoker` status based on count:* 
```{r}
table(insurance_df$smoker)
```

<br>

**2. Bivariate Analysis**

The figures below show the bivariate relationships of `charges` to `sex`, `region`, `smoker` and `children`.

```{r, fig.align='center', fig.width= 15, fig.height = 6}
p3_charges_sex <- ggplot(data = insurance_df, aes(x= sex, y = charges)) +
  geom_boxplot(alpha = 0.8, color = "dodgerblue3") + 
  scale_fill_brewer(palette= "Blues") +
  scale_y_continuous(label = dollar_format())+
  labs(x = "Sex",
       y = "Charges", 
       title = "Charges and Sex") +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "none")

p5_charges_region <- ggplot(data = insurance_df, aes(x= region, y = charges)) +
  geom_boxplot(alpha = 0.8, color = "dodgerblue3")+
  scale_fill_brewer(palette= "Blues") +
  scale_y_continuous(label = dollar_format())+
  labs(x = "Region",
       y = "Charges", 
       title = "Charges and Region") +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "none")

ggarrange(p3_charges_sex, p5_charges_region, ncol=2)

```

<br>

```{r, fig.align='center', fig.width=15, fig.height = 6}
p4_charges_smoker <- ggplot(data = insurance_df, aes(x= smoker, y = charges)) +
  geom_boxplot(alpha = 0.8, color = "dodgerblue3")+
  scale_y_continuous(label = dollar_format())+
  labs(x = "Smoker status",
       y = "Charges", 
       title = "Charges and Smoker status") +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "none")

p6_charges_children_num <- ggplot(data = insurance_df, aes(x= children_num, y = charges)) +
  geom_boxplot(alpha = 0.8, color = "dodgerblue3")+
  scale_y_continuous(label = dollar_format())+
  labs(x = "Number of Children",
       y = "Charges", 
       title = "Charges and Number of children") +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "none")

ggarrange(p4_charges_smoker, p6_charges_children_num, ncol=2)

```
<br>

The plot of `charges` against `smoker` clearly shows that the median charges for smokers are higher than non-smokers, and that the IQR are also larger for smokers. In terms of `region`, the median `charges` are almost same across all four regions, though the IQR or dispersion is highest for southeast. The same is observed in the plot of `charges` against `sex` where median charges is same between females and males though data on males have larger IQR. For the plot of `charges` against `children`, the median charges was observed to be increasing from having one  child to having four children, while median charges for those without children is quite higher and lower for those with five children.   

<br>

**3. Multivariate Analysis**

The figures below show the multivariate relationships of `charges`, `age`, `sex`, `smoker` and `bmi`.

```{r, fig.align="center", fig.width=15, fig.height = 6}

p1_charges_age_sex <- ggplot(data=insurance_df, aes(x= age, y = charges, color = sex)) +
  geom_point(size=2)+
  geom_smooth(method = "lm", aes(color = sex), size = 0.5)+
  scale_y_continuous(label = dollar_format())+
  scale_color_manual(values= c("goldenrod3", "dodgerblue3")) +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "bottom", legend.title = element_text(size = 15), legend.text = element_text(size = 15)) +
  labs(x = "Age",
       y = "Charges",
       col = "Sex",
       title = "Charges based on Age and Sex")

p2_charges_age_smoker <- ggplot(data=insurance_df, aes(x= age, y = charges, color = smoker)) +
  geom_point(size=2)+
  geom_smooth(method = "lm", aes(color = smoker), size = 0.5)+
  scale_y_continuous(label = dollar_format())+
  scale_color_manual(values= c("goldenrod3", "dodgerblue3")) +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "bottom", legend.title = element_text(size = 15), legend.text = element_text(size = 15)) +
  labs(x = "Age",
       y = "Charges",
       col = "Smoker",
       title = "Charges based on Age and Smoker status")

ggarrange(p1_charges_age_sex, p2_charges_age_smoker, ncol=2)

```
<br>

```{r, fig.align="center", fig.width=15, fig.height = 6}
p3_charges_bmi_sex <- ggplot(data=insurance_df, aes(x= bmi, y = charges, color = sex)) +
  geom_point(size=2)+
  geom_smooth(formula = 'y ~ x', method = "lm", aes(color = sex), linewidth = 1)+
  scale_color_manual(values= c("goldenrod3", "dodgerblue3")) +
  scale_y_continuous(label = dollar_format())+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "bottom", legend.title = element_text(size = 15), legend.text = element_text(size = 15)) +
  labs(x = "BMI",
       y = "Charges",
       col = "Sex",
       title = "Charges based on BMI and Sex")

p4_charges_bmi_smoker <- ggplot(data=insurance_df, aes(x= bmi, y = charges, color = smoker)) +
  geom_point(size=2)+
  geom_smooth(formula = 'y ~ x', method = "lm", aes(color = smoker), linewidth = 1)+
  scale_y_continuous(label = dollar_format())+
  scale_color_manual(values= c("goldenrod3", "dodgerblue3")) +
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  theme(legend.position = "bottom", legend.title = element_text(size = 15), legend.text = element_text(size = 15)) +
  labs(x = "BMI",
       y = "Charges",
       col = "Smoker",
       title = "Charges based on BMI and Smoker")

ggarrange(p3_charges_bmi_sex, p4_charges_bmi_smoker, ncol = 2)
```
<br>

The plots above show positive correlation between `charges` and `age` and between `charges` and `bmi`. It is also evident from the plots that there is not much difference in the correlation in terms of `sex` as compared to the correlation difference of `smoker`. In the plot of `charges` against `bmi` and `smoker`, the slope of the positive correlation line for smokers is steeper than the line for non-smoker at increasing BMI. This suggests possible interaction effect of `smoker` and `bmi`.  


<br>

The figure below shows the **correlation** among the variables. Clearly, `smoker` (yes) has the highest positive correlation with `charges`, followed by `age` and `bmi`.


```{r, fig.align='center', fig.width= 6, fig.height = 6}

model.matrix(~0+., data = insurance_df_og) %>% 
  cor(use = "pairwise.complete.obs") %>% 
  ggcorrplot(show.diag = F, type = "lower", lab = TRUE, lab_size = 3.5, colors = c("dodgerblue3", "white", "goldenrod3"),
             ggtheme= ggplot2::theme_gray, tl.cex = 10)

```
<br>

### III. Evaluation of Linear Regression Models

**A. Splitting the dataset**

The dataset was split into training data set (80%) and test data set (20%). The training data set was used to build the linear regression model and the test data set was used to evaluate the performance of the final model. 

```{r, echo=TRUE}
RNGkind(sample.kind = "Rounding")
set.seed(42)

train_index <-  sample(nrow(insurance_df), nrow(insurance_df)*0.8)

training_data <- insurance_df[train_index, ]
test_data <- insurance_df[-train_index, ]

nrow(training_data)
nrow(test_data)
```

<br>

**B. Linear regression models**

For this analysis, 4 models were evaluated, starting from a general simple model to an improved regression model.

1. Model 1 - General model including all independent variables
2. Model 2 - With added non-linear relationship
3. Model 3 - With added transformed variable
4. Model 4 - With added interaction relationship

Model results are detailed below. 

<br>

**Model 1 - General model including all independent variables**

This model fits a linear regression model that relates all 6 independent variables with the target `charges` variable (medical expenses).

<div align = 'center'> _charges = $\beta_{0}$ + $\beta_{1}$age + $\beta_{2}$bmi + $\beta_{3}$children + $\beta_{4}$sex + $\beta_{5}$smoker + $\beta_{6}$region + $\epsilon$_ </div>  

```{r, echo = TRUE}
model_01 <- lm(charges ~ age + bmi + children + sex + smoker + region, data = training_data)

summary(model_01)

```

<br>

The 'estimate' values in above model results refer to the estimated beta coefficients which indicate the increase in `charges` for an increase in each of the independent variable when the other variables are held constant. For instance, assuming everything else is equal, an additional child will result to an average of \$511 additional medical expenses (`charges`) per year. Furthermore, smokers have an average of \$24,177 more in medical expenses compared to non-smokers. 

<br>


**Model 2 - With added non-linear relationship**

```{r}
insurance_df$age2 <- insurance_df$age^2
training_data$age2 <- training_data$age^2
test_data$age2 <- test_data$age^2
```

This model includes a higher order term added to the regression Model 1 to account for a possible nonlinear relationship of the target variable to one of the independent variables. In this instance, the effect of `age` on `charges` may not be constant in all ages and that expenses may be disproportionately higher for older patients. Hence, a variable `age2` with values equivalent to `age` squared was created.

<div align = 'center'> _charges = $\beta_{0}$ + $\beta_{1}$age + $\beta_{2}$age^{2} + $\beta_{3}$bmi + $\beta_{4}$children + $\beta_{5}$sex + $\beta_{6}$smoker + $\beta_{7}$region + $\epsilon$_ </div>  

```{r, echo = TRUE}
model_02 <- lm(charges ~ age + age2 + bmi + children + sex + smoker + region, data = training_data)

summary(model_02)
```
<br> 

The result of this model shows that `charges` is more significantly related to `age2` than `age`.

<br>


**Model 3 - With added transformed variable**

```{r}
insurance_df$bmi30 <- ifelse(insurance_df$bmi >= 30, 1, 0)
training_data$bmi30 <- ifelse(training_data$bmi >= 30, 1, 0)
test_data$bmi30 <- ifelse(test_data$bmi >= 30, 1, 0)
```

This model also considers the possibility of `bmi` as not having a cumulative effect, but when BMI value is over a specific threshold value. This implies possibly no effect on medical expenses when patient is with normal weight but may have strong effect when patient is obese, i.e., when BMI is 30 or above. Hence, a binary indicator variable `bmi30` was created which has a value of 1 for `bmi` of at least 30, and 0 otherwise. 

<div align = 'center'> _charges = $\beta_{0}$ + $\beta_{1}$age + $\beta_{2}$age^{2} + $\beta_{3}$bmi + $\beta_{4}$bmi30 +$\beta_{5}$children + $\beta_{6}$sex + $\beta_{7}$smoker + $\beta_{8}$region + $\epsilon$_ </div>  

```{r, echo = TRUE}
model_03 <- lm(charges ~ age + age2 + bmi + bmi30 + children + sex + smoker + region, data = training_data)

summary(model_03)
```

<br>
The result shows that both `bmi` and `bmi30` have significant effect on medical expenses.

<br>


**Model 4 - With added interaction relationship**

This model considers the possibility of an interaction or combined effect of the variables `bmi30` and `smoker` (obesity and smoking) on the medical expenses. Hence, an interaction form `bmi30:smoker` was added.

<div align = 'center'> _charges = $\beta_{0}$ + $\beta_{1}$age + $\beta_{2}$age^{2} + $\beta_{3}$bmi + $\beta_{4}$bmi30 + $\beta_{5}$children + $\beta_{6}$sex + $\beta_{7}$smoker + $\beta_{8}$bmi30:smoker + $\beta_{9}$region + $\epsilon$_ </div>  

```{r, echo = TRUE}
model_04 <- lm(charges ~ age + age2 + bmi + bmi30 + children + sex + smoker + bmi30:smoker + region, data = training_data)

summary(model_04)
```

<br>
The result shows a significant effect on medical expenses of combined obesity and smoking.   

<br>


**C. Comparing the models**

```{r, include = TRUE}
model_list <- list(model_01, model_02, model_03, model_04)

func1 <- function(x){summary(x)$adj.r.squared}
func2 <- function(x){summary(x)$sigma}

model_results <- data.frame(
  models = c("model_01", "model_02", "model_03", "model_04"),
  r2 = unlist(lapply(model_list, func1)),
  rse = unlist(lapply(model_list, func2))
)

model_results
```

The four models were compared in terms of the 'residual standard error (RSE)' and the 'R-squared' values to determine which model has the better fit for the dataset .

Since RSE measures the standard deviation of the residuals (i.e., observed value minus the predicted value)in the regression model, the model with the smaller RSE is the better regression model that fit the dataset. On the other hand, the multiple R-squared value, called coefficient of determination, indicates how well a model as a whole explains the values of the dependent variable. The closer the value of this R-squared to 1.0, the better the model explains the data. However, since the four models have different numbers of explanatory variables, the 'adjusted R-squared value' were used for comparison. The adjusted R-squared corrects the multiple R-squared with penalty based on number of variables in the model.   

Based on the figures below, **Model 4** has the smallest RSE and highest adjusted R-squared, and hence, is the regression model that better fits the dataset. 


```{r, fig.align="center", fig.width= 15, fig.height = 6}
plot1_models <- ggplot(data = model_results, aes(x= models, y = r2, label = signif(r2, 3))) +
  geom_col(width= 0.7, color = "goldenrod1", fill = "goldenrod3")+
  geom_text(position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  labs(x = "R Squared",
       y = "Model",
       title = "Comparison of Model R squared values")

plot2_models <- ggplot(data = model_results, aes(x= models, y = rse, label = signif(rse, 3))) +
  geom_col(width= 0.7, color = "dodgerblue1", fill = "dodgerblue3")+
  geom_text(position = position_stack(vjust=0.5),
            colour = "white", size = 5)+
  theme(plot.title = element_text(size = 16), axis.title = element_text(size = 16), axis.text = element_text(size = 16))+
  labs(x = "RSE",
       y = "Model",
       title = "Comparison of Model RSE values")

ggarrange(plot2_models, plot1_models, ncol = 2)
```

<br>



**D. Predicting from test data**

```{r}
test_data$tst_yhat <- round(predict(model_04, newdata = test_data), 2)

```

The performance of Model 4 on new data was evaluated using the test data. Predicted  values of `charges` for the test data were obtained and plotted against known `charges` and shown in the figure below. The plot shows a good predictive performance to medical expenses around \$20,000 but slightly overpredict on `charges` above this. There are also few points showing large deviation of predicted from known values of `charges` in the test data. 

```{r, fig.align='center', fig.width=6.5, fig.height=5}
ggplot(data=test_data, aes(x=charges, y=tst_yhat)) +
  geom_point(color = "dodgerblue3", size=1.5) +
  geom_smooth(method = "lm", color = "goldenrod2", se= FALSE, linetype= "dashed", size = 0.5) +
  scale_y_continuous(label=dollar_format()) +
  scale_x_continuous(label=dollar_format()) +
  labs(x = "Known Charges",
       y = "Predicted charges",
       title= "Predicted against known outcomes plot")

```

<br>
<br>

### IV. Conclusion

In this project, predictive models for determining medical expenses based on certain patient characteristics were developed and compared. The selected model below, Model 4, which is an improved model including an added non-linear relationship (`age2`), a transformed variable (`bmi30`) and an interaction relationship (`bmi30:smoker`) was determined to have the better fit to the dataset among the models, evident from its smallest RSE (4490) and highest adjusted RSE (0.87). 

```{r, echo = TRUE}
model_04 <- lm(charges ~ age + age2 + bmi + bmi30 + children + sex + smoker + bmi30:smoker + region, data = training_data)
```

Based on the summary of Model 4, the variables with the greatest predictive power are `age2`, `children`, `smokeryes` and `bmi30:smokeryes`. A child or dependent has an averaged increased cost of \$708 in medical expenses per year. Furthermore, a smoker has an average increased cost of \$13,500 per year, and an obese smoker will have an additional \$19,894 per year in medical expenses. 

Although the model has good R-squared result, the plot showing predicted `charges` against known `charges` of the test data shows that the regression model does not perfectly fit the dataset. As the dataset seemed not large enough, adding more data (observations) to the dataset might improve the model and increase accuracy of the model in forecasting medical expenses. Moreover, further investigation and action on the outliers could also help in building more fitting models. 

<br>
<br>
<br>
